---
title: "Linux Distribution Popularity"
author: "Jared Prins"
date: "May 18, 2017"
output:
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r, echo=FALSE, include=FALSE}
library(dplyr)
library(plotly)
library(readODS)

distros <- read_ods(path = "distrowatch_data.ods", sheet = 2, col_names = TRUE, col_types = NULL,
                    na = "", skip = 0, formula_as_formula = FALSE, range = NULL)

hdr <- c("Rank", "Distribution", "HitsPerDay", "Year")

names(distros) <- hdr

ubuntu <- distros %>%
  filter(Distribution %in% c("Ubuntu", "Mint", "Debian") )
```


## Distrowatch

[Distrowatch](http://distrowatch.com/) is a website which provides news, popularity rankings, and other general information about various Linux distributions ([link][1]). It has ranked the popularity of linux distributions dating as far back as 2002.  

__It's important to note that this data is purely for fun.__ Even the creator of Distrowatch is quoted for saying that the data on Distrowatch is highly questionable.  


## 3 Popular Distributions

- Debian Linux has a long and proud history, being one of the earliest operating systems based on the Linux kernel ([link][2]). Many Linux distributions use Debian as a codebase. Two of the more popular ones are Ubuntu and Mint.
- Ubuntu Linux is named after the Southern African philosophy of ubuntu which can be loosely translated as "humanity to others" or "I am what I am because of who we all are" ([link][3]).
- Linux Mint is a community-driven Linux distribution based on Debian and Ubuntu that strives to be a "modern, elegant and comfortable operating system ([link][4]).


## History of Ubuntu, Mint and Debian Linux popularity
```{r plot, echo=FALSE}
p <- plot_ly(ubuntu, x = ~Year, y = ~HitsPerDay, text = ~Distribution, type = 'scatter', mode = 'markers', size = ~HitsPerDay, color = ~Distribution, colors = 'Paired',
             marker = list(opacity = 0.7, sizemode = 'diameter')) %>%
  layout(title = 'According to Distrowatch.com data',
         xaxis = list(showgrid = FALSE),
         yaxis = list(showgrid = FALSE),
         showlegend = FALSE)
p

```

## Brief Explanation

Debian has always had a loyal and slow growing user adoption.  But it was not a very user friendly Linux Distribution for everyday users.

Ubuntu changed this by adding user-friendly features on top of Debian. It drew massive user adoption.

Mint came out of nowhere in an attempt to be even more user friendly, particularly drawing on MS Windows users. Mint popularity exploded.

As more and more choices of linux distributions are available, even the popular distributions such as Mint and Ubuntu begin to lose market share. 


[1]: https://en.wikipedia.org/wiki/DistroWatch
[2]: https://en.wikipedia.org/wiki/Debian
[3]: https://en.wikipedia.org/wiki/Ubuntu_(operating_system)
[4]: https://en.wikipedia.org/wiki/Linux_Mint

