---
title: "DataProduct_Assignment1"
author: "Jared Prins"
date: May 16, 2017
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

setwd("E:\\_Personal\\Developing Data Products\\Assignment1")
rds.filename <- 'Property_Assessment_Data.rds'
censusfilename <- "census.rds"

Property_Assessment_Data <- readRDS(rds.filename)
dfCensus <- readRDS(censusfilename)


#filter out condos
library(dplyr)

df <- Property_Assessment_Data %>%
  filter(Assessment_Class == "Residential"  ) %>% 
  select(Neighbourhood, Assessed_Value_Dollar, Assessed_Value_Amount, Latitude, Longitude) %>%
  filter(Assessed_Value_Amount > 0) %>%
  group_by(Neighbourhood) %>%
  slice(which.max(Assessed_Value_Amount))


#Join by Neighbourhood
joined <- inner_join(df, dfCensus, by="Neighbourhood")

#icon_location <- "/home/jared/Documents/Coursera/Developing Data Products/Assignment1/house-icon3.png"
icon_location <- "house-icon3.png"

library(leaflet)

propertyIcon <- makeIcon(
  iconUrl = icon_location,
  iconWidth = 20,
  iconHeight = 20
)
```

## City of Edmonton Property Assessments Data Visualizations

The following four charts are for illustration only, for comparing 4 visualizations that map Property Assessment data in the City of Edmonton.

### Data

Using data as of December 31, 2017.

Assessment data comes from Edmonton's open data portal:
https://data.edmonton.ca/City-Administration/Property-Assessment-Data/q7d6-ambg

Edmonton census data (Population by Neighbourhood) is also used:
https://data.edmonton.ca/Census/2016-Census-Population-by-Age-Range-Neighbourhood-/phd4-y42v


### Code
All code can be found here: 

* Github master repo: https://github.com/petbadger/datasciencecoursera

The HTML Leaflet output is here:

* Github pages: http://petbadger.github.io/datasciencecoursera/index.html


## Map 1: Basic

A visualization showing only the geolocated highest single valued property for each residential neighbourhood.


```{r map1}
my_map1 <- leaflet(df) %>%
  addTiles() %>%
  addMarkers(lng=~Longitude, lat=~Latitude, popup=~Neighbourhood, label=~Assessed_Value_Dollar,
             icon=propertyIcon )
my_map1
```



## Map 2: Clustering

The same as Map 1, but properties are clustered for easier viewing. 

```{r map2}  
my_map2 <- leaflet(df) %>%
  addTiles() %>%
  addMarkers(lng=~Longitude, lat=~Latitude, popup=~Neighbourhood, label=~Assessed_Value_Dollar,
             clusterOptions = markerClusterOptions(), icon=propertyIcon )
my_map2
```


## Map 3: Bubble size

The same as Map 1, but size of bubble represents the value of the property, propotional to each other.

```{r map3}
my_map3 <- leaflet(joined) %>%
  addTiles() %>%
  addCircleMarkers(lng=~Longitude, lat=~Latitude, popup=~Neighbourhood, label=~Assessed_Value_Dollar,
                   weight=1, radius=~Assessed_Value_Amount/1000000 )
my_map3
```


## Map 4: Cluster and bubble

This map uses both clustering and bubble size feature. There is one small difference, the size of bubble in this case represents the size of the neighbourhood population. _(Note that the Census appears incomplete as there is 0 population for some neighbourhoods.)_

```{r map4}
#Size of bubble is proportional to Per 1000 population within the neighbourhood
my_map4 <- leaflet(joined) %>%
  addTiles() %>%
  addCircleMarkers(lng=~Longitude, lat=~Latitude, popup=~Neighbourhood, label=~Assessed_Value_Dollar,
                   clusterOptions = markerClusterOptions(), weight=1, radius=~Total_Pop/100 )
my_map4

```

